### 2014-04-09 paper/designspace.tex
paper mod start

\section{The Design Space of Pipes}\tovi{might want to italicize all words that are part of the design space, such as threadable, tree, star, etc}
\bjoern{would be nice to have a bridge from related work to here.}\tovi{tried to do this}
Our review of prior art indicates the potential benifits of using interior pipes for adding interactivity to 3D printed objects, although no general purpose tools currently exist for this purpose. To guide the design of such tool, we now proivde a deisgn space of the types of pipes that a maker may want to utlize. We have identified four relevant dimensions of pipe design: openings, path constraints \bjoern{note the rename - replicate in figure}\tovi{like the new name}, topologies, and inserted media.  We describe the space below.

\begin{figure}[t]
\centering
    \includegraphics[width=1.0\columnwidth]{figures/tubespace.pdf}
\caption{The design space of pipes.  See text for details. \bjoern{needs full caption.} \bjoern{The caption font is messed up. This may be from including some math package. Look up how to redefine the font to match the UIST template.}}
\label{fig:pipespace}
\end{figure}

>>>>

\section{The Design Space of Pipes}\tovi{might want to italicize all words that are part of the design space, such as threadable, tree, star, etc}
Our review of prior art indicates the potential benifits of using interior pipes for adding interactivity to 3D printed objects, although no general purpose tools currently exist for this purpose. To guide the design of such tool, we now proivde a deisgn space of the types of pipes that a maker may want to utlize. We have identified four relevant dimensions of pipe design: openings, path constraints, topologies, and inserted media.  We describe the space below.

\begin{figure}[t]
\centering
    \includegraphics[width=1.0\columnwidth]{figures/tubespace.pdf}
\caption{The design space of pipes.  See text for details. \bjoern{needs full caption.}}
\label{fig:pipespace}
\end{figure}

### 2014-04-09 paper/introduction.tex
paper mod start

\section{Introduction}
Makers, as well as professional designers, increasingly leverage 3D printers as tools for design work.  A wide array of objects, ranging from toy figurines to video game controllers and jet engine blades, are now prototyped (and some times even manufactured) using these machines.  Most models fabricated on 3D printers today are passive: they express the form, but not the function, of an object.

@@

We describe a new design space of pipes and hollow chambers for the purpose of interaction design, where variables include openings, topologies, and inserted media. By exploring this design space new opportunities for adding interactivity to 3D printed objects can be realized. For example, injecting conductive paint can turn pipes into conductors, which can in turn be used to integrate electronic components into prints. Electroluminescent (EL) wire can be threaded through pipes to create computer-controlled visual output. And air can be pumped through pipes to produce haptic effects.  Existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye}, Swept Frequency Capacitive Sensing \cite{Sato-touche}, and Jamming User Interfaces \cite{Follmer-jamming}, can be enhanced by the use of pipes: makers utilizing such I/O strategies can locate input and output at arbitrary points on a printed 3D object's surface.  \tovi{Is enhanced the right word? Or would "enabled" be better - since these techniques can now be applied to 3D printed objects} We leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of I/O to arbitrary locations on a device.

>>>>

\section{Introduction}
Makers, researchers, and professional designers (jointly "makers" for the remainder of this paper) increasingly leverage 3D printers as tools for design work.  A wide array of objects, ranging from toy figurines to video game controllers and jet engine blades, are now prototyped (and some times even manufactured) using these machines.  Most models fabricated on 3D printers today are passive: they express the form, but not the function, of an object.

@@

We describe a new design space of pipes and hollow chambers for the purpose of interaction design, where variables include openings, topologies, and inserted media. By exploring this design space new opportunities for adding interactivity to 3D printed objects can be realized. For example, injecting conductive paint can turn pipes into conductors, which can in turn be used to integrate electronic components into prints. Electroluminescent (EL) wire can be threaded through pipes to create computer-controlled visual output. And air can be pumped through pipes to produce haptic effects.  Existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye}, Swept Frequency Capacitive Sensing \cite{Sato-touche}, and Jamming User Interfaces \cite{Follmer-jamming}, can be enhanced by the use of pipes: makers utilizing such I/O strategies can locate input and output at arbitrary points on a printed 3D object's surface.  \tovi{Is enhanced the right word? Or would "enabled" be better - since these techniques can now be applied to 3D printed objects} \valkyrie{I like enhanced; we are allowing them to be applied in new places}  We leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of I/O to arbitrary locations on a device.

### 2014-04-09 paper/relatedwork.tex
paper mod start

Previous work has begun to investigate how to fabricate interactive objects using particular sensing and display technologies. Printed Optics (\cite{Willis-printedoptics}) uses light pipes to create integrated displays and sensed input optically through components added at print time; while Slyper et al., fabricate flexible robot armatures capable of sensing interaction via changes in air pressure \cite{Slyper-pressure}. Our work is inspired by these projects, and contributes a general design tool for modelling objects that can have such capabilities.

@@

\subsubsection{Fabrication with Printed Electronics}
Researchers are also investigating how to print conductors and place electronic components as part of additive manufacturing.  Sells, et al., \cite{Sells-reprap} use a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Sarik, et al., \cite{Sarik-tracebrush} use aerosol jetting of conductive material onto printed objects.  Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature} inject liquid metals. Our approach does not deposit such materials automatically, but creates structures inside a 3D model that make manual post-print injection of conductive ink possible. 
%Navarette, et al., created a functioning GPS routed in true 3D (not multilayer boards' 2.5D), hand-built using a one-off routing.  \bjoern{need more context for this}
%Our work differs from these because we offer \emph{algorithms and techniques} for \emph{internal}, \emph{true 3D} routing.
\tovi{should we say somewhere that 3d print manufacturers are considering conductive materials, but such printers would be extremely expensive?}

>>>>

Previous work has begun to investigate how to fabricate interactive objects using particular sensing and display technologies. Printed Optics \cite{Willis-printedoptics} uses light pipes to create integrated displays and sensed input optically through components added at print time; while Slyper et al., fabricate flexible robot armatures capable of sensing interaction via changes in air pressure \cite{Slyper-pressure}. Our work is inspired by these projects, and contributes a general design tool for modelling objects that can have such capabilities.

@@

\subsubsection{Fabrication with Printed Electronics}
Researchers are also investigating how to print conductors and place electronic components as part of additive manufacturing.  Sells, et al., \cite{Sells-reprap} use a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Other experiments have been done by the RepRap project\footnote{\url{http://reprap.org}} using plastic filament blended with conductive materials, however the reistance is too high to be useful in circuitry.  Sarik, et al., \cite{Sarik-tracebrush} use aerosol jetting of conductive material onto printed objects.  Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature} inject liquid metals.   Our approach does not deposit such materials automatically, but creates structures inside a 3D model that make manual post-print injection of conductive ink possible.
